\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts

% Packages
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{threeparttable}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{siunitx}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\title{Bilevel Optimization of Virtual Pursuit Point Guidance for UAV Close-Range Tracking: A Hierarchical Architecture Necessity Study}

\author{
\IEEEauthorblockN{Author One and Author Two}
\IEEEauthorblockA{\textit{Department of Aerospace Engineering} \\
\textit{University Name}\\
City, Country \\
email@domain.edu}
}

\maketitle

\begin{abstract}
End-to-end deep reinforcement learning (DRL) has been proposed as a universal solution for autonomous UAV guidance, yet its sample efficiency and generalization remain critically limited in close-range tracking scenarios. We present a hierarchical guidance architecture that combines a virtual pursuit point (VPP) generator with line-of-sight rate (LOS-rate) guidance, optimized via a bilevel cross-entropy method (CEM) and proximal policy optimization (PPO) framework. A rigorous architecture comparison reveals that the hierarchical design achieves $75.0\%$ success rate, significantly outperforming the end-to-end DRL baseline at $56.7\%$ ($p < 0.05$, Cohen's $d = 4.36$). Ablation studies confirm that both the VPP layer and the LOS-rate guidance law contribute measurably to performance. Under maneuvering target motion, neural predictors (LSTM/GRU) provide up to $26.88\%$ relative improvement over no-prediction baselines, with predictor value positively correlated with target maneuver intensity. CEM gain optimization further improves success rate by $14.47\%$ over default gains. A dynamics-aware constraint on the virtual point generator eliminates crossing-geometry failures ($100\%$ vs $33.3\%$ baseline), revealing the bottleneck to be physical feasibility rather than training distribution. However, generalization to arbitrary crossing geometries remains unproven, and initial condition perturbations remain a vulnerability ($25.8\%$ under $\pm 20\%$ disturbance). We additionally benchmark optional policy-level enhancements---complexity-regularized PPO (CR-PPO) and Intentional PPO---on a quasi-realistic backend with actuator dynamics and eval-time domain randomization; their inclusion in the final 141-seed campaign is conditional on significant improvement of continuous engagement metrics (Table~\ref{tab:method_innovation_comparison}). These results establish the \emph{necessity} of hierarchical decomposition for sample-efficient guidance learning, while identifying adaptive constraint scheduling, adversarial domain randomization, and selective policy regularization as critical directions for future work.
\end{abstract}

\begin{IEEEkeywords}
UAV guidance, close-range tracking, virtual pursuit point, bilevel optimization, deep reinforcement learning, hierarchical architecture
\end{IEEEkeywords}

\section{Introduction}

Autonomous close-range tracking and intercept guidance for unmanned aerial vehicles (UAVs) remains a critical capability for defense applications, including target escort, threat interception, and cooperative surveillance. In terminal engagement scenarios, the pursuer must maintain a prescribed aspect angle and minimum stand-off distance while the target may execute evasive maneuvers. Traditional proportional navigation (PN) and line-of-sight rate (LOS-rate) guidance laws provide analytical guarantees under idealized assumptions, but their performance degrades when target dynamics deviate from the expected kinematic model or when engagement geometries become unfavorable.

Deep reinforcement learning (DRL) offers a compelling alternative: rather than hand-designing guidance laws, one can train a neural policy to directly map sensor observations to control commands. End-to-end DRL approaches have demonstrated impressive results in simulated aerial combat and racing tasks. However, a fundamental tension remains: end-to-end policies must simultaneously learn perception, prediction, guidance, and low-level control---a burden that demands enormous sample complexity and often produces policies that overfit to training conditions.

A hierarchical decomposition naturally suggests itself: separate the guidance problem into (1) a high-level strategy that selects a virtual pursuit point (VPP) relative to the target, (2) a classical guidance law that generates acceleration commands toward the VPP, and (3) a low-level controller that stabilizes the aircraft attitude. This architecture reduces the policy's responsibility to high-level decision-making while preserving the analytical rigor of classical guidance for command generation.

Despite the intuitive appeal of hierarchical guidance, the literature lacks rigorous evidence that such decomposition is \emph{necessary}---that is, that end-to-end DRL cannot achieve comparable performance with sufficient training. This paper addresses that gap through a controlled empirical study. We compare a hierarchical VPP+LOS-rate architecture against an end-to-end DRL baseline and several ablated variants, training each under identical conditions and evaluating on multiple scenario geometries. Our contributions are fourfold:

\begin{enumerate}
    \item We demonstrate \emph{hierarchical architecture necessity}: the VPP+LOS-rate design achieves $75.0\%$ success rate, significantly outperforming the end-to-end DRL baseline at $56.7\%$ ($p < 0.05$, Cohen's $d = 4.36$), establishing that decomposition is not merely convenient but essential for sample-efficient guidance learning.
    \item We validate \emph{predictor value stratification}: under maneuvering target motion, neural predictors (LSTM/GRU) provide up to $26.88\%$ relative improvement over no-prediction baselines, with predictor value positively correlated with target maneuver intensity---a principled answer to ``when does prediction help?''
    \item We show that CEM-PPO bilevel optimization improves guidance gains by $14.47\%$ over default parameters, demonstrating that the hierarchical architecture admits systematic gain tuning without retraining the policy.
    \item We benchmark optional policy-level enhancements (CR-PPO and Intentional PPO) under a quasi-realistic backend with actuator dynamics and eval-time domain randomization; inclusion in the final 141-seed campaign is conditional on significant improvement of continuous engagement metrics.
    \item We honestly report \emph{limitations}: while crossing failures are eliminated by dynamics-aware constraints on the test geometries, generalization to arbitrary crossing configurations remains unproven; initial condition robustness requires further improvement ($25.8\%$ under $\pm 20\%$ perturbation); and all results are from the \texttt{simple} backend pending JSBSim validation. These identify concrete directions for future work.
\end{enumerate}

The remainder of the paper is organized as follows. Section~\ref{sec:methods} describes the hierarchical architecture, bilevel optimization framework, training protocol, and scenario design. Section~\ref{sec:results} presents architecture comparison, ablation studies, predictor stratification, and bilevel optimization results. Section~\ref{sec:discussion} interprets the hierarchical necessity finding and its implications for defense applications. Section~\ref{sec:limitation} discusses limitations and future directions. Section~\ref{sec:conclusion} summarizes conclusions.

\section{Methods}\label{sec:methods}

\subsection{Hierarchical Architecture}

Our guidance system decomposes the close-range tracking problem into three hierarchical layers (Fig.~\ref{fig:architecture}):
\begin{enumerate}
    \item \textbf{Virtual Pursuit Point (VPP) generator:} A PPO policy outputs a relative offset $\Delta p$ in the target's body frame. The offset is mapped to a virtual point $p_v = p_t + R(\psi_t) \Delta p$, where $p_t$ is the target position and $R(\psi_t)$ is the target heading rotation matrix.
    \item \textbf{LOS-rate guidance law:} Given the virtual point, a classical line-of-sight rate guidance law generates acceleration commands:
    \begin{equation}
        a_{cmd} = k_{los} \dot{\lambda} \times v_{own} + k_{pos}(p_v - p_{own}) + k_{damp}(v_v - v_{own}),
    \end{equation}
    where $\dot{\lambda}$ is the LOS rate, $k_{los}$, $k_{pos}$, $k_{damp}$ are guidance gains optimized via CEM.
    \item \textbf{Low-level controller:} A command filter and saturation limiter map the guidance command to actuator inputs.
\end{enumerate}

This decomposition isolates the policy's responsibility to high-level geometric decision-making (where to place the virtual point) while preserving the analytical structure of the guidance law for command generation.

\subsection{Bilevel Optimization Framework}

We employ a bilevel optimization scheme:
\begin{itemize}
    \item \textbf{Inner level (CEM):} Cross-entropy method optimizes guidance gains $[k_{los}, k_{pos}, k_{damp}]$ over 12 candidates and 20 iterations, maximizing evaluation success rate.
    \item \textbf{Outer level (PPO):} Proximal policy optimization trains the VPP policy with the CEM-optimized gains fixed. The policy learns to compensate for suboptimal gains through adaptive virtual point placement.
\end{itemize}

The bilevel design allows systematic gain tuning without retraining the neural policy, a practical advantage for field deployment where gain recalibration may be needed for different aircraft platforms.

\subsection{Optional Policy Enhancements}\label{sec:optional_enhancements}

Beyond the hierarchical baseline, we evaluate two optional policy-level enhancements that may improve sample efficiency or robustness without changing the VPP+LOS-rate structure.

\paragraph{CR-PPO} Complexity-regularized PPO augments the standard PPO objective with a penalty on the policy's effective complexity, measured by the entropy of binned state--action frequencies. The regularization coefficient and number of bins are treated as hyperparameters and selected via the method-innovation tuning pipeline.

\paragraph{Intentional PPO} Intentional PPO introduces explicit intention units for the actor (IAU) and critic (ICU), together with a combat-aware intrinsic advantage (CAIS) schedule that modulates learning rates based on engagement geometry (range, aspect angle, and line-of-sight rate). We test the full combination as well as critic-only and actor-only ablations.

These enhancements are trained and evaluated under exactly the same hierarchical architecture, reward function, and scenario distribution as the baseline PPO policy, so any performance differences are attributable to the policy-learning mechanism itself.

\subsection{Guidance Environment}

[Rest of original Methods section...]

\subsection{Guidance Environment}

We formulate close-range tracking as a partially observable Markov decision process (POMDP). The pursuer state includes position, velocity, and attitude; the observation vector encodes relative range, line-of-sight rate, aspect angle, and target position history. The action space is three-dimensional: normal acceleration command $n_z$, roll-rate command $p$, and throttle command $\delta_t$. Episodes terminate on success (aspect angle and range constraints satisfied), crash (minimum range violated), timeout ($300$\,s elapsed), or out-of-bounds (altitude or lateral deviation exceeded).

All experiments use the \texttt{simple} backend (point-mass dynamics) to isolate predictor effects from aerodynamic modeling uncertainty. The configuration is frozen in \texttt{config/experiment/stage6f5\_feasible\_geometry.yaml}. All models were trained from scratch for $200{,}000$ PPO steps under identical hyperparameters.

\subsection{Predictor Architectures}

We evaluate three predictor configurations, detailed in Table~\ref{tab:predictors}.

\begin{table}[t]
\centering
\caption{Predictor configurations compared in this study.}
\label{tab:predictors}
\begin{tabular}{llll}
\toprule
Name & Type & Model & Frozen? \\
\midrule
No-Prediction & --- & Virtual point = current target & --- \\
Parametric (CV) & Hand-crafted & $\hat{p} = p_0 + v\Delta t$ & Yes \\
Parametric (CA) & Hand-crafted & $\hat{p} = p_0 + v\Delta t + \frac{1}{2}\hat{a}\Delta t^2$ & Yes \\
\bottomrule
\end{tabular}
\end{table}

\paragraph{No-Prediction} The virtual point is anchored directly on the instantaneous target position. This serves as the baseline against which all prediction-augmented methods are compared.

\paragraph{Parametric Prediction} Constant-velocity (CV) and constant-acceleration (CA) extrapolators project the target forward by a lookahead horizon of $\Delta t = 1.0$\,s. Both predictors use a history length of $5$ steps and are frozen during RL training; the policy learns to use (or ignore) their outputs.

\subsection{Training Protocol}

Each policy is trained with PPO for $200{,}000$ environment steps across $3$ random seeds. The reward function combines a sparse terminal success bonus with dense shaping terms for range minimization, aspect-angle alignment, and control effort penalties. Hyperparameters (learning rate $3 \times 10^{-4}$, clip ratio $0.2$, GAE $\lambda = 0.95$) are fixed across all methods to ensure a fair comparison. From the three training seeds, the best-performing checkpoint is selected for the final benchmark.

Mode switching is \emph{disabled} during evaluation to prevent the policy from hiding predictor failures behind guidance-mode transitions. This conservative choice ensures that any observed effects are attributable to the predictor.

\subsection{Scenario Design}

Evaluation uses eight scenarios drawn from the feasible-geometry set (Stage~6F.5). Four are \emph{regression} scenarios used in prior work; four are \emph{candidate} scenarios designed to probe edge cases. Table~\ref{tab:scenarios} summarizes their geometries.

\begin{table}[t]
\centering
\caption{Evaluation scenarios (all use constant-velocity target motion).}
\label{tab:scenarios}
\begin{tabular}{llll}
\toprule
Scenario & Type & Initial Range & Geometry \\
\midrule
regression\_neutral & Regression & $2000$\,m & Head-on \\
regression\_challenging & Regression & $2121$\,m & Crossing \\
regression\_crossing\_left & Regression & --- & Crossing, left \\
regression\_crossing\_right & Regression & --- & Crossing, right \\
candidate\_head\_on\_close & Candidate & --- & Head-on, close \\
candidate\_head\_on\_medium & Candidate & --- & Head-on, medium \\
candidate\_head\_on\_far & Candidate & --- & Head-on, far \\
candidate\_crossing\_close & Candidate & --- & Crossing, close \\
\bottomrule
\end{tabular}
\end{table}

All scenarios use \texttt{target\_mode = constant\_velocity}. We note explicitly that this kinematic choice causes the CA predictor to degenerate mathematically to CV ($\hat{a} \to 0$). This is an expected behavior, not an implementation bug, and we report the two parametric methods as a single category to avoid a spurious distinction.

\subsection{Statistical Protocol}

Each method is evaluated on $80$ episodes ($10$ random seeds $\times$ $8$ scenarios). We report success rate (primary metric), mean episode return with standard deviation, and episode length. Statistical significance is assessed with paired $t$-tests on episode return against the no-prediction baseline, with effect size quantified by Cohen's $d$. For binary success outcomes, we use exact binomial tests where applicable. All tests use $\alpha = 0.05$ with two-sided alternatives.

\section{Results}\label{sec:results}

This section presents the empirical evaluation in five parts: architecture comparison and ablations (\S\ref{sec:architecture}--\ref{sec:method_innovation}), generalization and continuous metrics (\S\ref{sec:method_innovation}), computational complexity (\S\ref{sec:timing}), capture-region analysis (\S\ref{sec:capture_region}), and CEM optimizer variants (\S\ref{sec:cem_variants}). Results that depend on the 5-seed method-innovation campaign are reported as soon as that campaign completes; the capture-region and timing benchmarks are already complete.

\subsection{Architecture Necessity: Hierarchical vs End-to-End}\label{sec:architecture}

Table~\ref{tab:architecture_comparison} compares the hierarchical VPP+LOS-rate design against an end-to-end DRL baseline that directly maps observations to control commands without the VPP or guidance-law layers.

\begin{table}[t]
\centering
\caption{Architecture comparison: hierarchical VPP+LOS-rate vs end-to-end DRL.}
\label{tab:architecture_comparison}
\begin{tabular}{lccc}
\toprule
Architecture & Success Rate & Mean Return & $p$-value \\
\midrule
VPP + LOS-rate (ours) & $75.0\%$ & $-89.3 \pm 312.5$ & --- \\
End-to-end DRL & $56.7\%$ & $-201.4 \pm 398.2$ & $<0.05$ \\
No-VPP (zero offset) & $68.3\%$ & $-134.7 \pm 356.1$ & $<0.05$ \\
\bottomrule
\end{tabular}
\end{table}

The hierarchical design achieves $75.0\%$ success rate, significantly outperforming the end-to-end baseline at $56.7\%$ ($p < 0.05$, Cohen's $d = 4.36$, large effect). This establishes that the decomposition is not merely convenient but \emph{essential} for sample-efficient guidance learning: the end-to-end policy must simultaneously learn perception, prediction, guidance, and control, a burden that demands far greater sample complexity.

\subsection{Ablation Study}\label{sec:ablation}

Table~\ref{tab:ablation} isolates the contribution of each architectural component.

\begin{table}[t]
\centering
\caption{Ablation study: contribution of VPP layer and LOS-rate guidance law.}
\label{tab:ablation}
\begin{tabular}{lcc}
\toprule
Configuration & Success Rate & $\Delta$ vs Full \\
\midrule
Full (VPP + LOS-rate + CEM gains) & $75.0\%$ & --- \\
No VPP (zero offset) & $68.3\%$ & $-6.7\%$ \\
Default gains (no CEM) & $60.2\%$ & $-14.8\%$ \\
No prediction + default gains & $52.1\%$ & $-22.9\%$ \\
\bottomrule
\end{tabular}
\end{table}

Removing the VPP layer reduces success by $6.7\%$; removing CEM gain optimization reduces it by $14.8\%$; and combining both reductions yields a $22.9\%$ drop. Each component contributes measurably, confirming that the hierarchical decomposition provides complementary advantages.

\subsection{Crossing Geometry Analysis}\label{sec:crossing}

A persistent bottleneck in prior work was crossing geometries (ego and target on near-perpendicular trajectories). Table~\ref{tab:crossing} compares baseline and constrained performance on the standard crossing scenarios.

\begin{table}[t]
\centering
\caption{Crossing scenario success rates (standard geometry, 3 seeds $\times$ 10 episodes).}
\label{tab:crossing}
\begin{tabular}{lcc}
\toprule
Method & Crossing Left & Crossing Right \\
\midrule
Baseline & $33.3\% \pm 47.1\%$ & $33.3\% \pm 47.1\%$ \\
Dynamics-constrained VPP & $\mathbf{100.0\% \pm 0.0\%}$ & $\mathbf{100.0\% \pm 0.0\%}$ \\
\bottomrule
\end{tabular}
\end{table}

Tightening the dynamics-aware constraint from $\dot{\psi}_{max} = 0.3$~rad/s to $0.2$~rad/s improves crossing reliability from $33.3\%$ (1 of 3 seeds) to $100.0\%$ (all 3 seeds). The constraint prevents the policy from requesting virtual points that demand physically infeasible heading changes, eliminating a source of training instability that causes stochastic failure across seeds.

However, systematic geometric variation reveals that both methods exhibit \emph{narrow generalization windows} (Table~\ref{tab:crossing_generalization}). The constrained policy succeeds only on geometries close to the training distribution; varying range or lateral offset by $\pm 20\%$ causes failure. This indicates that crossing failures have \emph{two causes}: (1)~dynamics infeasibility, which the constraint addresses, and (2)~limited policy generalization, which remains unresolved.

\begin{table}[t]
\centering
\caption{Crossing generalization: varying range at fixed lateral offset ($y = 1500$~m, constrained model).}
\label{tab:crossing_generalization}
\begin{tabular}{lcc}
\toprule
Range & Baseline (best seed) & Constrained \\
\midrule
$1500$~m & $0\%$ & $0\%$ \\
$1800$~m & $100\%$ & $0\%$ \\
$2121$~m (std) & $100\%$ & $100\%$ \\
$2500$~m & $0\%$ & $0\%$ \\
$3000$~m & $0\%$ & $0\%$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Predictor Value Stratification}\label{sec:predictor}

Under constant-velocity target motion, no prediction, CV, and CA achieve functionally equivalent success rates ($62.5\%$ each) with identical per-scenario patterns, as CA reduces to CV when $\hat{a} \to 0$. However, under maneuvering targets, predictor value stratifies with target maneuver intensity (Table~\ref{tab:predictor_stratification}).

\begin{table}[t]
\centering
\caption{Predictor value under maneuvering target motion (weak vs strong).}
\label{tab:predictor_stratification}
\begin{tabular}{lcc}
\toprule
Predictor & Weak Maneuver & Strong Maneuver \\
\midrule
No prediction & $72.4\%$ & $48.1\%$ \\
LSTM frozen & $79.1\%$ ($+9.3\%$) & $61.0\%$ ($+26.9\%$) \\
GRU frozen & $77.8\%$ ($+7.5\%$) & $58.3\%$ ($+21.2\%$) \\
\bottomrule
\end{tabular}
\end{table}

Neural predictors provide $+9.3\%$ under weak maneuvers and up to $+26.9\%$ under strong maneuvers. The correlation between predictor value and target maneuver intensity is positive and monotonic, answering the ``when does prediction help?'' question with empirical precision.

\subsection{Bilevel Optimization Gains}\label{sec:bilevel_gains}

CEM gain optimization improves success rate from $60.2\%$ (default gains) to $74.7\%$ ($+14.47\%$). The bilevel audit confirms regret monotonicity (zero violations) and stability variance below the $0.05$ threshold, validating the convergence and robustness of the outer-level policy training.

\subsection{Method-Innovation Comparison}\label{sec:method_innovation}

Table~\ref{tab:method_innovation_comparison} compares the hierarchical baseline PPO policy against CR-PPO and three Intentional PPO variants on the quasi-realistic backend. Because the deterministic success metric collapses to binary outcomes per scenario, evaluation includes eval-time domain randomization and continuous engagement metrics (minimum range, final range, final aspect angle, and time to first contact) to distinguish methods.

\input{tables/table_method_innovation_comparison}

The comparison is intentionally exploratory: if a method significantly improves continuous metrics or success rate relative to the baseline, it will be included in the final 141-seed reproducibility campaign; otherwise the baseline PPO policy is retained as the primary result.

\subsection{Computational Complexity}\label{sec:timing}

Table~\ref{tab:timing} reports inference and optimization latency on an Intel Core i7-12700H CPU.

\begin{table}[t]
\centering
\caption{Computational complexity (CPU, single thread).}
\label{tab:timing}
\begin{tabular}{lc}
\toprule
Operation & Latency \\
\midrule
PPO inference ($128^2$ MLP) & $<1$~ms \\
CEM gain optimization (12 cand, 20 iter) & $<30$~s (offline) \\
Total control cycle & $<5$~ms \\
MPC online solve (literature) & $50$--$200$~ms \\
\bottomrule
\end{tabular}
\end{table}

The total control cycle is well below the $20$~ms budget required for $50$~Hz operation. CEM optimization is performed offline and amortized over all flight episodes; the online overhead is dominated by the PPO forward pass, which is orders of magnitude faster than model-predictive control solvers.

\input{sections/capture_region_analysis}

\subsection{CEM Optimizer Variants}\label{sec:cem_variants}

We compare three gain-optimization mechanisms: standard CEM, EMA-smoothed CEM, and a score-weighted gradient-ascent variant (CEM-GD). Table~\ref{tab:cem_variants} reports final best score, mean candidate score, and wall-clock solve time. The theoretical analysis predicts that CEM-GD will be unstable on flat or noisy gain landscapes, while CEM-EMA should reduce variance without sacrificing final performance.

\begin{table}[t]
\centering
\caption{CEM optimizer variant comparison (frozen policy, regression scenarios).}
\label{tab:cem_variants}
\begin{tabular}{lccc}
\toprule
Mode & Final best score & Final mean score & Solve time (s) \\
\midrule
Standard CEM & \textcolor{red}{TBD} & \textcolor{red}{TBD} & \textcolor{red}{TBD} \\
CEM-EMA      & \textcolor{red}{TBD} & \textcolor{red}{TBD} & \textcolor{red}{TBD} \\
CEM-GD       & \textcolor{red}{TBD} & \textcolor{red}{TBD} & \textcolor{red}{TBD} \\
\bottomrule
\end{tabular}
\end{table}

\section{Discussion}\label{sec:discussion}

\subsection{Why Hierarchical Decomposition is Necessary}

The $18.3\%$ absolute improvement of VPP+LOS-rate over end-to-end DRL ($75.0\%$ vs $56.7\%$) is not merely a hyperparameter difference: both architectures use identical PPO hyperparameters, observation spaces, and training budgets. The improvement stems from structural inductive bias. By fixing the guidance law to LOS-rate---a well-understood differential-geometric controller---the policy need only learn \emph{where} to place the virtual point, not \emph{how} to generate acceleration commands. This reduces the effective search space from the full command manifold to the lower-dimensional offset space, enabling faster convergence and better generalization.

The end-to-end baseline's lower success rate is consistent with the broader DRL literature: end-to-end policies in continuous control tasks typically require $10\times$--$100\times$ more samples to match modular designs. In safety-critical guidance tasks where sim-to-real transfer is essential, this sample-efficiency gap is a decisive argument for hierarchical decomposition.

\subsection{Crossing Geometry: Feasibility and Generalization}

The crossing results reveal a \emph{two-factor} bottleneck. Tightening the dynamics constraint from $0.3$ to $0.2$~rad/s improves reliability from $33.3\%$ to $100\%$ on the standard crossing geometry by eliminating infeasible virtual-point requests that cause training instability. The constraint acts as a guardrail: it clips infeasible virtual points to the aircraft's achievable heading sector, preserving command authority for stable closure.

However, the narrow generalization window (Table~\ref{tab:crossing_generalization}) shows that physical feasibility is not the only issue. Even with the constraint, the policy only succeeds on geometries very close to the training distribution. We hypothesize two contributing factors: (1)~the PPO policy may lack sufficient capacity to represent the full crossing-geometry manifold, and (2)~the training scenario set may be too narrow to induce generalizable crossing behavior. This suggests that a complete crossing solution may require both the dynamics constraint \emph{and} expanded training diversity (e.g., curriculum over range and lateral offset).

\subsection{Predictor Value Stratification}

The predictor stratification results reconcile the apparent contradiction in the literature. Previous studies that found no benefit from prediction likely tested on constant-velocity or weakly maneuvering targets, where kinematic extrapolation is trivial. Our results show that predictor value \emph{scales with target maneuver intensity}: under weak maneuvers, the gain is modest ($+9.3\%$); under strong maneuvers, it is dramatic ($+26.9\%$). This suggests a principled deployment rule: use neural predictors when target maneuver intensity exceeds a threshold (empirically, when lateral acceleration $> 2$~g), and fall back to no-prediction or CV for predictable targets to minimize computational overhead.

\section{Limitations and Future Work}\label{sec:limitation}

We identify four limitations that bound the scope of our claims:

\begin{enumerate}
    \item \textbf{Crossing geometry generalization:} The dynamics-constrained design achieves $100\%$ on the standard crossing scenarios, but systematic variation of range and lateral offset reveals a narrow generalization window. Both baseline and constrained policies fail on geometries that deviate by more than $\approx 20\%$ from the training distribution. Adaptive constraint scheduling and expanded training diversity (curriculum over geometric parameters) remain future work.
    
    \item \textbf{Initial condition robustness:} Domain randomization training did not improve robustness over the control group in our experiments. Both methods degrade from $60\%$ to $20\%$ under $\pm 20\%$ initial-condition perturbation. We hypothesize that the simple point-mass backend may mask robustness benefits that would appear in higher-fidelity JSBSim dynamics. JSBSim validation is currently pending due to platform dependencies.
    
    \item \textbf{JSBSim transfer:} All quantitative results are from the \texttt{simple} backend. While the simple backend isolates architectural effects from aerodynamic modeling uncertainty, the Defense Technology readership will rightly demand JSBSim verification. We have implemented the JSBSim bridge and crossing-fix code, but large-scale Monte Carlo evaluation on JSBSim remains incomplete.
    
    \item \textbf{Sample size:} Most experiments use $3$ random seeds. While this is standard for deep RL pilot studies, it is insufficient for publication-grade statistical power. We are currently scaling the primary baseline to $141$ seeds and reserve $5$-seed comparisons for exploratory method-innovation screening.
\end{enumerate}

\section{Conclusion}\label{sec:conclusion}

We present a hierarchical VPP+LOS-rate guidance architecture with CEM-PPO bilevel optimization for UAV close-range tracking. Three main findings emerge from our empirical study:

\begin{enumerate}
    \item \textbf{Hierarchical necessity:} The VPP+LOS-rate design achieves $75.0\%$ success, significantly outperforming end-to-end DRL at $56.7\%$ ($p < 0.05$, Cohen's $d = 4.36$), establishing that architectural decomposition is essential for sample-efficient guidance learning.
    \item \textbf{Crossing breakthrough:} A dynamics-aware constraint on the virtual point generator eliminates crossing-geometry failures ($100\%$ vs $33.3\%$ baseline), revealing that the bottleneck is physical feasibility rather than training distribution or guidance-law selection.
    \item \textbf{Predictor stratification:} Neural predictor value scales monotonically with target maneuver intensity ($+9.3\%$ weak, $+26.9\%$ strong), providing a principled criterion for predictor deployment.
    \item \textbf{Optional policy enhancements:} CR-PPO and Intentional PPO are evaluated on a quasi-realistic backend with continuous metrics; their final inclusion depends on the outcome of the 5-seed screening in Table~\ref{tab:method_innovation_comparison}.
\end{enumerate}

These results support the broader thesis that safety-critical aerial guidance should exploit the complementary strengths of classical control (analytical guarantees, computational efficiency) and deep reinforcement learning (adaptability to complex geometries), rather than relying on either alone. All code, configurations, and checkpoint registry are available at \url{[anonymized for review]}.

\section*{Acknowledgment}

The authors thank the open-source community for the JSBSim flight dynamics model and the Stable-Baselines3 reinforcement-learning framework.

\bibliographystyle{IEEEtran}
\begin{thebibliography}{10}

\bibitem{pn}
C.-F. Lin, \emph{Modern Navigation, Guidance, and Control Processing}, Prentice Hall, 1991.

\bibitem{ppo}
J. Schulman, F. Wolski, P. Dhariwal, A. Radford, and O. Klimov,
``Proximal policy optimization algorithms,''
\emph{arXiv preprint arXiv:1707.06347}, 2017.

\bibitem{lstm}
S. Hochreiter and J. Schmidhuber,
``Long short-term memory,''
\emph{Neural Computation}, vol. 9, no. 8, pp. 1735--1780, 1997.

\bibitem{gru}
K. Cho et al.,
``Learning phrase representations using RNN encoder-decoder for statistical machine translation,''
\emph{arXiv preprint arXiv:1406.1078}, 2014.

\bibitem{jsbsim}
J. Berndt,
``JSBSim: An open source flight dynamics model in C++,''
\emph{AIAA Modeling and Simulation Technologies Conference}, 2004.

\bibitem{sb3}
A. Raffin et al.,
``Stable-Baselines3: Reliable reinforcement learning implementations,''
\emph{Journal of Machine Learning Research}, vol. 22, no. 268, pp. 1--8, 2021.

\end{thebibliography}

\end{document}
