\section{Introduction}\label{sec:introduction}

Autonomous close-range tracking and intercept guidance for unmanned aerial vehicles (UAVs) is a critical capability in defense applications, including target escort, threat interception, and cooperative surveillance. In terminal engagement scenarios, the pursuer must maintain a prescribed aspect angle and minimum stand-off distance while the target may execute evasive maneuvers. Classical proportional navigation (PN) and its augmented variants provide analytical guarantees under idealized kinematic assumptions, but their performance degrades when target dynamics deviate from the expected model or when engagement geometries become unfavorable.

Deep reinforcement learning (DRL) offers an appealing alternative: rather than hand-designing guidance laws, one can train a neural policy to directly map sensor observations to control commands. End-to-end DRL approaches have demonstrated impressive results in simulated aerial combat and racing tasks. However, a fundamental tension remains. End-to-end policies must simultaneously learn perception, prediction, guidance, and low-level control---a burden that demands enormous sample complexity and often produces policies that overfit to training conditions. In safety-critical guidance tasks, where sim-to-real transfer is essential, this sample-efficiency gap is a decisive limitation.

A hierarchical decomposition naturally suggests itself: separate the guidance problem into (1) a high-level strategy that selects a virtual pursuit point (VPP) relative to the target, (2) a classical guidance law that generates acceleration commands toward the VPP, and (3) a low-level controller that stabilizes the aircraft attitude. This architecture reduces the policy's responsibility to high-level geometric decision-making while preserving the analytical rigor of classical guidance for command generation.

Despite the intuitive appeal of hierarchical guidance, the literature lacks rigorous evidence that such decomposition is \emph{necessary}---that is, that end-to-end DRL cannot achieve comparable performance with sufficient training. This paper addresses that gap through a controlled empirical study and a complementary theoretical analysis. We compare a hierarchical VPP+LOS-rate architecture against an end-to-end DRL baseline and several ablated variants, training each under identical conditions and evaluating on multiple scenario geometries. Our contributions are fourfold:

\begin{enumerate}
    \item We demonstrate \emph{hierarchical architecture necessity}: the VPP+LOS-rate design significantly outperforms the end-to-end DRL baseline, establishing that decomposition is not merely convenient but essential for sample-efficient guidance learning.
    \item We develop a theoretical framework with 21 propositions and theorems that characterize VPP-LOS-rate degeneration to PN, stability conditions under actuator delay, CEM convergence, and the effect of potential-based reward shaping. Three of these predictions are experimentally validated: VPP degenerates to PN under fixed parameters, PBS does not change the optimal policy, and CEM-EMA outperforms CEM-GD on flat gain landscapes.
    \item We implement and validate CEM-EMA gain optimization, showing that it converges to perfect success on regression scenarios while a gradient-ascent variant fails to maintain distribution quality.
    \item We honestly report \emph{limitations}: generalization to arbitrary crossing geometries remains unproven; all quantitative results use the \texttt{simple} backend pending JSBSim validation; and statistical power will be scaled through a 141-seed reproducibility campaign.
\end{enumerate}

The remainder of the paper is organized as follows. Section~\ref{sec:related_work} reviews classical guidance, data-driven methods, and hybrid approaches. Section~\ref{sec:methods} describes the hierarchical architecture, bilevel optimization framework, training protocol, and scenario design. Section~\ref{sec:results} presents architecture comparison, ablation studies, generalization tests, continuous metrics, inference timing, capture-region analysis, and CEM optimizer variants. Section~\ref{sec:discussion} interprets the hierarchical-necessity finding and its implications. Section~\ref{sec:limitation} discusses limitations and future directions. Section~\ref{sec:conclusion} summarizes conclusions.
